### 2019-02-20 paper/birth_spacing_lit.tex
Update spacing lit notes

Using the NLSY79 and NLSY79 Child and Young Adult Surveys we investigate 
the effect of the age difference between between siblings (spacing) on
educational achievement. Because spacing may be endogenous, we use an
instrumental variables strategy that exploits variation in spacing driven
by miscarriages. The IV results indicate that a one-year increases in
spacing increases test scores for older siblings by about 0.17 standard
deviations. These results are larger than the OLS estimates, suggesting
that failing to account for the endogeneity of spacing may understate
its benefits. For younger siblings, we find no causal impact of spacing
on test scores.

@@

this analysis unequivocally suggests that excess fe-
male mortality in India has both economic and
cultural roots and any explanation must con-
sider both economic and cultural factors.

@@

Examines preferred spacing as compared to observed spacing. In 5 of 20 countries the
actual spacing is significantly longer than the preferred spacing.

@@

This paper’s contribution is to directly estimate the causal relationship between family size and the desired sex ratio and to quantify an important cause of rising sex ratios. The challenge in estimating this effect is to isolate exogenous variation in the fertility level. The approach I use is to elicit sex composition preferences at different fertility levels: a hypothetical fertility level is specified to the survey respondent, and she is asked, given that fertility level, what is her preferred composition of boys and
girls. 5 By imposing the total number of children, one can characterize the respondents’ sex ratio preferences at different exogenously determined fertility levels.

>>>>

Using the NLSY79 and NLSY79 Child and Young Adult Surveys we investigate

the effect of the age difference between between siblings (spacing) on
educational achievement. Because spacing may be endogenous, we use an
instrumental variables strategy that exploits variation in spacing
driven
by miscarriages. The IV results indicate that a one-year increases in
spacing increases test scores for older siblings by about 0.17 standard
deviations. These results are larger than the OLS estimates, suggesting
that failing to account for the endogeneity of spacing may understate
its benefits. For younger siblings, we find no causal impact of spacing
on test scores.

@@

this analysis unequivocally suggests that excess female mortality in
India has both 
economic and cultural roots and any explanation must consider both
economic and cultural 
factors.

@@

Examines preferred spacing as compared to observed spacing. In 5 of 20
countries the
actual spacing is significantly shorter than the preferred spacing.

@@

This paper’s contribution is to directly estimate the causal
relationship between family size and the desired sex ratio and to
quantify an important cause of rising sex ratios. The challenge in
estimating this effect is to isolate exogenous variation in the
fertility level. The approach I use is to elicit sex composition
preferences at different fertility levels: a hypothetical fertility
level is specified to the survey respondent, and she is asked, given
that fertility level, what is her preferred composition of boys and
girls. By imposing the total number of children, one can characterize
the respondents’ sex ratio preferences at different exogenously
determined fertility levels.
